Zero-shot adaptation by meta-mapping between tasks

### 2019-09-04 writing/embedded_meta_learning_main.tex
Further tweaks, cutting a bit

\begin{document}
\title{Zero-shot task adaptation by meta-mapping from task to task}
\author{%
Andrew K. Lampinen\\
Department of Psychology\\
Stanford University\\
\texttt{lampinen@stanford.edu}\\
\And
James L. McClelland\\
Department of Psychology\\
Stanford University\\
\texttt{mcclelland@stanford.edu}\\
}
\date{}
\maketitle

@@

\vspace{-0.75em} % section
\section{Introduction}
\vspace{-0.5em} % section
Humans are able to use and reuse knowledge more flexibly than most deep learning models can \citep[e.g.][]{Lake2016, Marcus2018}. One fundamental reason for this is that humans are aware of what we are trying to compute and why. By contrast, there is a fundamental separation of knowledge within most deep neural networks -- although deep networks represent knowledge about data (in their activations) and knowledge about the structure of tasks (in their parameters), they do not represent any relationships between data and tasks. That is, a neural network's knowledge about what is being computed is only implicitly accessible to those computations. \par
\looseness=-1
There are a number of advantages to representing knowledge about data and tasks together. In particular, it can grant the ability to rapidly adapt behavior to a new task. The problem of rapid learning has been partially addressed by meta-learning systems \citep[e.g.][see also section \ref{sec_discussion}]{Santoro2016, Finn2017a, Finn2018, Stadie2018, Botvinick2019}. However, humans can use our knowledge of a task to flexibly alter our behavior in accordance with a change in task demands or a single instruction. For example, once we learn to play a game, we can immediately switch to playing in order to lose, and can achieve reasonable performance without any retraining (i.e. zero-shot). Deep learning systems at present generally lack this representational flexibility. \par
In this paper, we propose a new class of tasks based on this idea: meta-mappings, i.e. mappings between tasks (see below). This type of transfer is easily accessible to humans \citep{Lake2016}, but is generally inaccessible to most deep-learning models. To address this challenge, we propose using architectures which essentially take a functional perspective on meta-learning, and exploit the idea of homoiconicity. (A homoiconic programming language is one in which programs in the language can be manipulated by programs in the language, just as data can.) By treating both data and task behaviors as functions, we can conceptually think of both data \emph{and} learned task behaviors as transformable. This yields the ability to not only learn to solve new tasks, but to learn how to transform these solutions in response to changing task demands. We demonstrate that our architectures can flexibly remap their behavior to address the meta-mapping challenge. By allowing the network to recurse on its representations, our approach is able to achieve this flexibility parsimoniously, without adding additional parameters beyond what we use for basic meta-learning. We suggest that approaches like EML will be key to building more intelligent and flexible deep learning systems. \par

@@

\caption{The EML system can perform meta-mappings from language cues rather than meta-mapping examples, and can even generalize these. Compare to fig. \ref{meta_map_results}, which shows the same results when using examples instead of language. (\subref{poly_meta_map_results_language}) In the polynomials domain, language cues still lead to good performance, even on held-out tasks or held-out meta-mappings, although examples perform slightly better (fig. \ref{poly_meta_map_results_examples}). (\subref{cards_meta_map_results_language}) Similarly, in the cards domain, language cues perform well. Error-bars are bootstrap 95\%-CIs.}

@@

Language is a fundamental piece of human flexibility, and we would be remiss to ignore it. In many of the intuitive illustrations we've given, the examples of the meta-mapping are implicit in your prior knowledge about the world. For example, ``try to lose at go'' does not give explicit examples of the ``lose'' meta-mapping, but rather relies on prior knowledge of what losing means. This is a much more efficient way to achieve zero-shot performance on a new losing task. In order to replicate this, we trained the EML system with both meta-mappings based on examples, and meta-mappings based on language. In the language-based meta-mappings, a language input identifying the meta-mapping (but not the basic task to apply it to) is provided to the network. This language input is encoded by a language encoder, and then provided as input to $\mathcal{H}$ instead of the output of $\mathcal{M}$. The meta-mapping then proceeds as normal --- $\mathcal{H}$ parameterizes $F$, which is used to transform the embedding of the input task to produce an embedding for the target task. \par

@@

Our architecture builds directly off of prior work on HyperNetworks \citep{Ha2016} -- networks which parameterize other networks -- and other recent applications thereof, such as guessing parameters for a model to accelerate model search \citep[e.g.][]{Brock2018a, Zhang2019}, and meta-learning \citep[e.g.]{Li2019a, Rusu2019}. Our work is also related to the longer history of work on different time-scales of weight adaptation \citep{Hinton1982, Kumaran2016} that has more recently been applied to meta-learning contexts \citep[e.g.][]{Ba2016, Munkhdalai2017, Garnelo2018} and continual learning \citep[e.g.]{Hu2019}. It is more abstractly related to work on learning to propose architectures \citep[e.g.][]{Zoph2016, Cao2019}, and to models that learn to select and compose skills to apply to new tasks \citep[e.g.][]{Andreas, Andreas2016, Tessler2016, Reed2015, Chang2019a}. In particular, some of the work in domains like visual question answering has explicitly explored the idea of building a classifier conditioned on a question \citep{Andreas, Andreasa}, which is related to one of the possible computational paths through our architecture. Work in model-based reinforcement learning has also partly addressed how to transfer knowledge between different reward functions \citep[e.g.][]{Laroche2017}; our approach is more general. Indeed, our insights could be combined with model-based approaches, for example our approach could be used to adapt a task embedding, which would then be fed to a learned planning model to execute that behavior.\par 
There has also been other recent interest in task (or function) embeddings. Achille et al. \citep{Achille2019} recently proposed computing embeddings for visual tasks from the Fisher information of the parameters in a model partly tuned on the task. They show that this captures some interesting properties of the tasks, including some types of semantic relationships, and can help identify models that can perform well on a task. Rusu and colleagues recently suggested a similar meta-learning framework where latent codes are computed for a task which can be decoded to a distribution over parameters \citep{Rusu2019}. Other recent work has tried to learn representations for skills \citep[e.g.][]{Eysenbach2019} or tasks \citep[e.g.]{Hsu2019} for exploration and representation learning. Our perspective can be seen as a generalization of these that allows for remapping of behavior by meta-tasks. To the best of our knowledge none of the prior work has explored meta-mappings, and therefore it has not allowed for zero-shot performance of a new task. \par

@@

\paragraph{Future Directions:} We think that the general perspective of embedding tasks and considering meta-mappings will yield many fruitful future directions. We hope that our work will inspire more exploration of behavioral adaptation, in areas beyond the simple domains we considered here. To this end, we suggest the creation of meta-learning datasets which include information not only about tasks, but about the relationships between them. For example, as we alluded to above, tasks such as visual question-answering \citep[e.g.][]{Antol2015} can be usefully interpreted from this perspective (a question essentially yields a function from images to answers, and there are many systematic relationships between these functions), as can reinforcement learning tasks which involve executing instructions \citep[e.g.][]{Hermann2017, Co-Reyes2019}. Furthermore, we think our work provides a novel perspective on the types of flexibility that human intelligence exhibits, and thus hope that it may have implications for cognitive science. \par 
We do not necessarily believe that the particular architecture we have suggested is the best architecture for addressing these problems generally, although it has a number of desirable characteristics. However, because of the way the architecture is modularized, it is relatively easy to change any aspect of it without altering the others. (Note that we compare some variations in appendix \ref{app_lesion_results}.) For example, although we only considered task networks $F$ that are feed-forward and of a fixed depth, this could be replaced with a recurrent architecture to allow more adaptive computation, or even a more complex architecture \citep[e.g.][]{Reed2015, Graves2016}. Our work also opens the possibility of doing unsupervised learning over function representations for further learning, which relates to long-standing ideas in cognitive science about how humans represent knowledge \citep{Clark1993}. \par 

@@

In this paper, we've highlighted a new type of flexibility in the form of meta-mapping -- mapping between tasks. Meta-mapping can produce zero-shot performance on novel tasks, based on their relationship to old tasks. However, achieving this type of flexibility is difficult in architectures that separate computation from the knowledge of what is being computed. To address this, our EML architecture explicitly represents tasks by function embeddings, and derives the computation from these embeddings via a HyperNetwork. This allows for our key insight: using transformations over task embeddings to transform the system's behavior. This is analogous to how transformations over data perform basic tasks. Following this analogy, we've proposed a meta-learning system which takes a functional perspective, embeds tasks and data into a shared latent space, and then meta-learns to make basic- and meta-mappings over this space. The requisite meta-task sample complexity is small; we showed generalization to unseen meta-mappings after training on only 20 meta-mappings. That is, we are able to achieve zero-shot performance on a novel task based on a relationship between tasks that was never trained on. This is a step closer to human-level flexibility.\par

>>>>

\begin{document}
\title{Zero-shot adaptation by meta-mapping between tasks}
\author{%
Andrew K. Lampinen\\
Department of Psychology\\
Stanford University\\
\texttt{lampinen@stanford.edu}\\
\And
James L. McClelland\\
Department of Psychology\\
Stanford University\\
\texttt{mcclelland@stanford.edu}\\
}
\date{}
\maketitle

@@

\vspace{-0.75em} % section
\section{Introduction}
\vspace{-0.5em} % section
Humans are able to use and reuse knowledge more flexibly than most deep learning models can \citep{Lake2016, Marcus2018}. One fundamental reason for this is that humans are aware of what we are trying to compute and why. By contrast, there is a fundamental separation of knowledge within most deep neural networks -- although deep networks represent knowledge about data (in their activations) and knowledge about the structure of tasks (in their parameters), they do not represent any relationships between data and tasks. That is, a neural network's knowledge about what is being computed is only implicitly accessible to those computations. \par
\looseness=-1
There are a number of advantages to representing knowledge about data and tasks together. In particular, it can grant the ability to rapidly adapt behavior to a new task. The problem of rapid learning has been partially addressed by meta-learning systems \citep[see also section \ref{sec_discussion}]{Santoro2016, Finn2017a, Finn2018, Stadie2018, Botvinick2019}. However, humans can use our knowledge of a task to flexibly alter our behavior in accordance with a change in task demands or a single instruction. For example, once we learn to play a game, we can immediately switch to playing in order to lose, and can achieve reasonable performance without any retraining (i.e. zero-shot). Deep learning systems at present generally lack this representational flexibility. \par
In this paper, we propose a new class of tasks based on this idea: meta-mappings, i.e. mappings between tasks (see below). This type of transfer is easily accessible to humans \citep{Lake2016}, but is generally inaccessible to most deep-learning models. To address this challenge, we propose using architectures which essentially take a functional perspective on meta-learning, and exploit the idea of homoiconicity. (A homoiconic programming language is one in which programs in the language can be manipulated by programs in the language, just as data can.) By treating both data and task behaviors as functions, we can conceptually think of both data \emph{and} learned task behaviors as transformable. This yields the ability to not only learn to solve new tasks, but to learn how to transform these solutions in response to changing task demands. We demonstrate that our architectures can flexibly remap their behavior to address the meta-mapping challenge. By allowing the network to recurse on its representations, our approach is able to achieve this flexibility parsimoniously, without adding additional parameters beyond what we use for basic meta-learning. We suggest that approaches like EML will be key to building more intelligent and flexible deep learning systems. \par

@@

\caption{The EML system can perform meta-mappings from language cues rather than meta-mapping examples. Compare to fig. \ref{meta_map_results}, which shows the same results when using examples instead of language. (\subref{poly_meta_map_results_language}) In the polynomials domain, language cues still lead to good performance, even on held-out tasks or held-out meta-mappings, although examples perform slightly better (fig. \ref{poly_meta_map_results_examples}). (\subref{cards_meta_map_results_language}) Similarly, in the cards domain, language cues perform well. Error-bars are bootstrap 95\%-CIs.}

@@

Language is fundamental to human flexibility, and we would be remiss to ignore it. In many of the intuitive illustrations we've given, the examples of the meta-mapping are implicit in your prior knowledge about the world. For example, ``try to lose at go'' does not give explicit examples of the ``lose'' meta-mapping, but rather relies on prior knowledge of what losing means. This is a much more efficient way to achieve zero-shot performance on a new losing task. In order to replicate this, we trained the EML system with both meta-mappings based on examples, and meta-mappings based on language. In the language-based meta-mappings, a language input identifying the meta-mapping (but not the basic task to apply it to) is encoded by a language encoder, and then provided as the input to $\mathcal{H}$ (instead of an output from $\mathcal{M}$). The meta-mapping then proceeds as normal --- $\mathcal{H}$ parameterizes $F$, which is used to transform the embedding of the input task to produce an embedding for the target. \par

@@

Our architecture builds directly off of prior work on HyperNetworks \citep{Ha2016} -- networks which parameterize other networks -- and other recent applications thereof, such as guessing parameters for a model to accelerate model search \citep[e.g.][]{Brock2018a, Zhang2019}, and meta-learning \citep[e.g.]{Li2019a, Rusu2019}. Our work is also related to the longer history of work on different time-scales of weight adaptation \citep{Hinton1982, Kumaran2016} that has more recently been applied to meta-learning contexts \citep[e.g.][]{Ba2016, Munkhdalai2017, Garnelo2018} and continual learning \citep[e.g.]{Hu2019}. It is more abstractly related to work on learning to propose architectures \citep[e.g.][]{Zoph2016, Cao2019}, and to models that learn to select and compose skills to apply to new tasks \citep[e.g.][]{Andreas, Andreas2016, Tessler2016, Reed2015, Chang2019a}. In particular, some of the work in domains like visual question answering has explicitly explored the idea of building a classifier conditioned on a question \citep{Andreas, Andreasa}, which is related to one of the possible computational paths through our architecture. Work in model-based reinforcement learning has also partly addressed how to transfer knowledge between different reward functions \citep[e.g.][]{Laroche2017}; our approach is more general. Indeed, our insights could be combined with model-based approaches, for example our approach could be used to adapt a task embedding, which would then be used by a learned planning model.\par 
There has also been other recent interest in task (or function) embeddings. Achille et al. \citep{Achille2019} recently proposed computing embeddings for visual tasks from the Fisher information of the parameters in a model partly tuned on the task. They show that this captures some interesting properties of the tasks, including some types of semantic relationships, and can help identify models that can perform well on a task. Rusu and colleagues recently suggested a similar meta-learning framework where latent codes are computed for a task which can be decoded to a distribution over parameters \citep{Rusu2019}. Other recent work has tried to learn representations for skills \citep[e.g.][]{Eysenbach2019} or tasks \citep[e.g.]{Hsu2019} for exploration and representation learning. Our perspective can be seen as a generalization of these that allows for remapping of behavior via meta-tasks. To the best of our knowledge none of the prior work has explored zero-shot performance of a task via meta-mappings. \par

@@

\paragraph{Future Directions:} We think that the general perspective of embedding tasks and considering meta-mappings will yield many fruitful future directions. We hope that our work will inspire more exploration of behavioral adaptation, in areas beyond the simple domains we considered here. To this end, we suggest the creation of meta-learning datasets which include information not only about tasks, but about the relationships between them. For example,  as can reinforcement learning tasks which involve executing instructions \citep[e.g.][]{Hermann2017, Co-Reyes2019} can be usefully interpreted from this perspective. Furthermore, we think our work provides a novel perspective on the types of flexibility that human intelligence exhibits, and thus hope that it may have implications for cognitive science. \par 
We do not necessarily believe that the particular architecture we have suggested is the best architecture for addressing these problems, although it has a number of desirable characteristics. However, the modularization of the architecture makes it easy to modify. (We compare some variations in appendix \ref{app_lesion_results}.) For example, although we only considered task networks $F$ that are feed-forward and of a fixed depth, this could be replaced with a recurrent architecture to allow more adaptive computation, or even a more complex architecture \citep[e.g.][]{Reed2015, Graves2016}. Our work also opens the possibility of doing unsupervised learning over function representations for further learning, which relates to long-standing ideas in cognitive science about how humans represent knowledge \citep{Clark1993}. \par 

@@

We've highlighted a new type of flexibility in the form of meta-mapping -- mapping between tasks. Meta-mapping can produce zero-shot performance on novel tasks, based on their relationship to old tasks. However, achieving this type of flexibility is difficult in architectures that separate computation from the knowledge of what is being computed. To address this, our EML architecture explicitly represents tasks by function embeddings, and derives the computation from these embeddings via a HyperNetwork. This allows for our key insight: using transformations over task embeddings to transform the system's behavior. This is analogous to how transformations over data perform basic tasks. Following this analogy, the EML architecture takes a functional perspective, embeds tasks and data into a shared latent space, and then meta-learns to make basic- and meta-mappings over this space. This approach is parsiomious, because it uses the same networks for basic- and meta-mappings. Perhaps because of this, the requisite meta-task sample complexity is small; we showed generalization to unseen meta-mappings after training on only 20 meta-mappings. That is, we are able to achieve zero-shot performance on a novel task based on a relationship between tasks that was never trained on. This is a step closer to human-level flexibility.\par
